Blum axioms and nondeterministic computation of functions

### 2016-05-19 paper.tex
paper.tex: Moved two phrases between parentheses to footnotes.

One important theorem that can be proven using acceptable Gödel numberings alone
is the recursion theorem
\cite[p.~181]{Rogers1987}.
It states that, if $\phi$ is any acceptable Gödel numbering
and $f$ is any total recursive function,
then there is some program $w_0$ such that $\phi_{w_0} = \phi_{f(w_0)}$.
That is,
if $f$ is any systematic transformation on programs,
there is a program $w$ (a fixed point for $f$) whose meaning under $\phi$ is unchanged.
The recursion theorem can be used,
for example, to show the existence of quines
(programs whose output are their own source codes)
in any Turing-complete programming language:
choose $f$ to be the function that, given a program $w$,
returns another program $f(w)$ that prints the string $w$ when run.
(For Turing machines, for instance,
we can encode the bits of $w$ in the transition table of $f(w)$.)
Then, by the recursion theorem,
there is some program $w_0$ that is equivalent to its transformed version $f(w_0)$;
thus, $w_0$ already writes the string $w_0$, its own source code.
Therefore, any programming language has quines \cite[p.~227]{Kozen2006}.

@@

\begin{proposition}
    If there is a nondeterministic Turing machine
    that computes in polynomial time, according to Goldreich's definition,
    the characteristic function of the satisfiability problem,
    then $\NP = \coNP$.
\end{proposition}

($\NP = \coNP$ implies that $\NP = \PH$; that is,
the polynomial hierarchy collapses to the first level \cite[p.~280]{Kozen2006}.)

>>>>

One important theorem that can be proven using acceptable Gödel numberings alone
is the recursion theorem
\cite[p.~181]{Rogers1987}.
It states that, if $\phi$ is any acceptable Gödel numbering
and $f$ is any total recursive function,
then there is some program $w_0$ such that $\phi_{w_0} = \phi_{f(w_0)}$.
That is,
if $f$ is any systematic transformation on programs,
there is a program $w$ (a fixed point for $f$) whose meaning under $\phi$ is unchanged.
The recursion theorem can be used,
for example, to show the existence of quines
(programs whose output are their own source codes)
in any Turing-complete programming language:
choose $f$ to be the function that, given a program $w$,
returns another program $f(w)$ that prints the string $w$ when run.\footnote{
    For Turing machines, for instance,
    we can encode the bits of $w$ in the transition table of $f(w)$.
}
Then, by the recursion theorem,
there is some program $w_0$ that is equivalent to its transformed version $f(w_0)$;
thus, $w_0$ already writes the string $w_0$, its own source code.
Therefore, any programming language has quines \cite[p.~227]{Kozen2006}.

@@

\begin{proposition}
    If there is a nondeterministic Turing machine
    that computes in polynomial time, according to Goldreich's definition,
    the characteristic function of the satisfiability problem,
    then $\NP = \coNP$.\footnote{
        $\NP = \coNP$ implies that $\NP = \PH$; that is,
        the polynomial hierarchy collapses to the first level \cite[p.~280]{Kozen2006}.
    }
\end{proposition}
